Rotation effects on the Lyman-$\alpha$ line morphology in distant galaxies

### 2013-05-27 paper/Analisis.tex
two more paragraphs in conclusions

\begin{abstract}
Rotation is present in the gas kinematics of galaxies up to the
highest redshifts. In this paper we present for the first time
radiative transfer calculations that show the impact of rotation on
the morphology of the Lyman $\alpha$ line. To this end we construct
simplified models where a galaxy is modeled as an homogeneous sphere
composed as an homogenous mixture of dust and hydrogen at a constant
temperature. These spheres have an solid-body rotation with linear
velocities at the surface in the range $0-300$ \kms. We consider
radiation sources both in the center of the rotating cloud and also
homogeneously distributed around the sphere. We find that higher
rotational velocities increase the width of each peak in the outgoing
line profile while it also increases the amount of Lyman alpha photons
escaping in the line center. This trends makes that for high
rotational velocities and large Hydrogen optical depths the double
peak of the line tends to be erased an be replaced by a single peak the
lines center. This is more pronounced for radiation sources
homogeneously distributed. Concerning the escape fraction we find that
rotation does not have any effect, provided that all the sources are
centrally emitted. However in the case of homogeneously emitted
sources we measure an increase of about a factor of $2$ in the escape
fraction for higher rotational velocity vealues.

@@

Untill now we have shown the main effects of gas and dust rotation in the Ly$\alpha$ line morphology such as the escape fraction, the FWHM the maxima position, also we see that shape the outgoing spectra depends in the position of an esternl observer. It is important to see if rotational effects are detectable in observational methods involving the Ly$\alpha$ line. By far this is not and extensive study, it is just and ....

One of the most used methods to detect high redshift galaxies using the Ly$\alpha$ line is using a narrowband selection, for this reason we focus this analysis with the

We make this analysis based in the results obtained by Steidel (2011) in this work (EXPLAIN a lit of bit more about their work) they used tree narrowband filters for tree different redshifts resumed in Table~\ref{table:NBfilters}, We want to know how much the integrated flux change due to rotational effects in this NB filters, for the models we simulated with CLARA.

@@

In table \ref{table:NBfilters} we present the results of the flux in every narrowband filter for the homogenous model at different velocity and hydrogen optical depth $\tau_{0}$. As is well known an increase in the optical depth makes the line peacks separation bigger. In fact for some cases the line width is larger than the NB filer width fig.(put ref fig) for those cases we modify the redshift in order to make the peak maxima match the NB filter center. For all cases we found that as velocity increase the flux is less. \\
In the case of the dusty model we found the same trend of the flux with the velocity but in this cases the effect is no that strong. 
We also found some dependency ith the viewing angle.

@@

1. Width and maxima effects

2. Avg N of scatterings

3. Integrated flux.

>>>>

\begin{abstract}
Rotation is present in the gas kinematics of galaxies up to the
highest redshifts. In this paper we present for the first time
radiative transfer calculations that show the impact of rotation on
the morphology of the Lyman $\alpha$ line. To this end we construct
simplified models where a galaxy is modeled as an homogeneous sphere
composed as an homogenous mixture of dust and hydrogen at a constant
temperature. These spheres have a solid-body rotation with linear
velocities at the surface in the range $0-300$ \kms. We consider
radiation sources both in the center of the rotating cloud and also
homogeneously distributed around the sphere. We find that higher
rotational velocities increase the width of each peak in the outgoing
line profile while it also increases the amount of Lyman alpha photons
escaping in the line center. This trends makes that for high
rotational velocities and large Hydrogen optical depths the double
peak of the line tends to be erased an be replaced by a single peak the
lines center. This is more pronounced for radiation sources
homogeneously distributed. Concerning the escape fraction we find that
rotation does not have any effect, provided that all the sources are
centrally emitted. However in the case of homogeneously emitted
sources we measure an increase of about a factor of $2$ in the escape
fraction for higher rotational velocity vealues.

@@

Untill now we have shown the main effects of gas and dust rotation in the Ly$\alpha$ line morphology such as the escape fraction, the FWHM the maxima position, also we see that shape of the outgoing spectra depends on the position of the observer. It is important to see if rotational effects are detectable in observational methods involving the Ly$\alpha$ line.

One of the most used methods to detect high redshift galaxies using the Ly$\alpha$ line is using a narrowband selection, we make this analysis based on the results obtained by Steidel (2011) in this work (EXPLAIN a lit of bit more about their work) they used tree narrowband filters for tree different redshifts resumed in Table~\ref{table:NBfilters}, We want to know how much the integrated flux change due to rotational effects in this NB filters, for the models we simulated with CLARA.

@@

In table \ref{table:NBfilters} we present the results of the flux in every narrowband filter for the homogenous model at different velocities and hydrogen optical depth $\tau_{0}$. As is well known an increase in the optical depth makes the line peacks separation bigger. In fact for some cases the line width is larger than the NB filer width fig.(put ref fig) for those cases we modify the redshift in order to make the peak maxima match the NB filter center. For all cases we found that as velocity increase the flux is less. \\
In the case of the dusty model we found the same trend of the flux with the velocity but in this cases the effect is no that strong. 
We also found some dependency with the viewing angle.

@@

As first we see how the width of the line changes using a modified FWHM explained in section \ref{sec:widthpeak}, and we found that as gas bulk rotation incrase also the width increase in a factor of $2-3$ in comparison with the static case. we also take into account the influence of the observer position is and we found that for observers with a line of sight perpendicular to the axis of rotation measure a larger ($15\%$) line width than those aligned with the rotation axis.\\

As many observational Ly$\alpha$ (Kulas et al) spectra are double peaked  and those provide important information about the inflows/outflows of gas content, we study the effect of rotation in the position of this peaks and we find that the position of the maxima does not change with rotation.\\

One of the key features in the study of the Ly$\alpha$ line is the escape fraction, when rotational effects are taking into account we found that the escape fraction increase in about $20\%-30\%$ for the homogenous sphere model. (Nscatt)

4. Integrated flux.
